---
title: "Reproducible Research - Course Project 1"
author: "Aron O'Connor"
date: "August 22, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading and Preprocessing Data
Load the activity.csv file from the working directory:

```{r}
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

# Mean Steps Per Day
Calculate the total number of steps taken per day, make a histogram of the total 
number of steps taken per day, and report the mean and median of the total number
of steps taken per day:

```{r}
activity_agg <- aggregate(activity$steps ~ activity$date,activity,sum)
names(activity_agg) <- c("date","steps")
activity_agg
plot(activity_agg$date,activity_agg$steps,type="h",main="Steps per Day Histogram"
     ,xlab="Date",ylab="Steps")
mean(activity_agg$steps)
median(activity_agg$steps)

```

# Average Daily Activity Pattern
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and 
the average number of steps taken, averaged across all days (y-axis).  Determine
which interval contains the highest average number of steps

```{r}
activity_interval <- aggregate(activity$steps ~ activity$interval,activity,mean)
names(activity_interval) <- c("interval","steps")
plot(activity_interval$interval,activity_interval$steps,type="l",main="Average 
     Steps per Interval",xlab="Interval",ylab="Average Steps")
max_average <- max(activity_interval$steps)
activity_interval[which(activity_interval$steps == max_average),1]
```

# Dealing with Missing Values
Calculate the number of incomplete cases in the data set.
```{r}
sum(complete.cases(activity)=="FALSE")
```

Fill in the missing step values in the data set and create a new data frame with
no 'NA's called 'activity_complete'.
```{r}
activity_complete <- data.frame()
for (i in 1:17568){
        if (is.na(activity[i,1])){
                activity_complete[i,1] <- activity_interval[match(activity[i,3],
                activity_interval$interval),2]
        } else {
                activity_complete[i,1] <- activity[i,1]
        }
}
activity_complete <- cbind(activity_complete,activity[,2:3])
names(activity_complete) <- c("steps","date","interval")
```
Plot a histogram of the total number of steps taken per day and calculate the 
mean and median of the new data set.
```{r}
activity_agg_complete <- aggregate(activity_complete$steps ~ activity_complete$date,
                        activity_complete,sum)
names(activity_agg_complete) <- c("date","steps")
plot(activity_agg_complete$date,activity_agg_complete$steps,type="h",main="Steps
     per Day Histogram",xlab="Date",ylab="Steps")
mean(activity_agg_complete$steps)
median(activity_agg_complete$steps)
```
The more complete data is surprisingly close to the original set!  By filling
in the 'NA' data, the median and mean are now the same number.

# Weekends vs. Weekdays
Add a factor variable to the data set that indicates whether the day is a 
weekday or weekend.
```{r}
weekdays1 <- c("Monday","Tuesday","Wednesday","Thursday","Friday")
activity_complete$weekday <- factor((weekdays(activity_complete$date) %in% weekdays1),
                                levels=c(TRUE,FALSE),labels=c("weekend","weekday"))
```
Create a panel plot containing a time series plot (i.e. type = "l") of the 5-minute 
interval (x-axis) and the average number of steps taken, averaged across all 
weekdays or weekends (y-axis).
```{r}

```
